ST563 Final Project

### 2021-03-11 ST563-FP.Rmd
added code for boosting added code for LASSO

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model
library(glmnet) # fitting LASSO

@@

  * `caret`: to do the heavy lifting of training and tuning the models
  * `tidyverse`: for all the data reading and wrangling
  * `knitr`: for rmarkdown table outputs
  * `rmarkdown`: for output documents
  * `corrplot`: for correlation structure visualization
  * `leaps`: for subset selection
  * `rpart`: fitting the basic classification tree
  * `rpart.plot`: visualization of the classification tree
  * `randomForest`: fitting random forest models
  * `kernlab`: fitting the support vector machine
  * `class`: fitting the k-nearest neighbor model

@@

```{r}
# Get X matrix
X <- model.matrix(quality ~ ., data = wine)[, -1]

# Fit LASSO model and plot result
lm.lasso <- glmnet::cv.glmnet(X, wine$quality, alpha = 1)
lambda <- lm.lasso$lambda.min
plot(lm.lasso)

# Fit LASSO model with best tune
lasso.best <- glmnet(X, wine$quality, alpha = 1)
predict(lasso.best, s = lambda, type = "coefficients")

# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
smre <- lm.sub %>% summary()

@@

# Best subset selection
best.betas <- coef(lm.sub, id = 6) %>% round(2)
best.betas
```

@@

# Print out regression tree visualization
tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
rpart.plot(tree.reg.out, main="Regression Tree")

# Print out classification tree visualization
tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
rpart.plot(tree.class.out, main="Classification Tree")

# Print out confusion matrix
list("RMSE" = tree.reg.rmse,
     "Class Results" = tree.class.res)
```

@@

# Obtain RMSE and results
rf.reg.pred <- predict(rf.reg, newdata = tst)
rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)

@@

# Print out results in a list
list("RMSE" = rf.reg.rmse,
     "Class Results" = rf.class.res)
```

@@

# Print out results in a list
list("Results" = svm.class.res,
     "Confusion Matrix" = svm.class.CM$table)
```

@@

# Obtain RMSE and results
knn.reg.pred <- predict(knn.reg, newdata = tst)
knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)

@@

# Print out results in a list
list("RMSE" = knn.reg.rmse,
     "Class Results" = knn.class.res)
```

@@

# Print out results in a list
list("Results" = glm.class.res,
     "Confusion Matrix" = glm.class.CM$table)
```

## LASSO Regression
LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero due to a penalty. This is different from traditional shrinkage methods since the form of the LASSO penalty actually forces some of the coefficients to exactly zero. This helps achieve variable selection and yields sparser models, similar to subset selection. In particular, the LASSO estimate minimizes:
$$
\sum_{i=1}^n\left( y_i - \beta_0 - \sum_{j=1}^p\beta_jx_{ij}\right)^2 + \lambda \sum_{j=1}^p|\beta_j| = RSS + \lambda \sum_{j=1}^p|\beta_j|
$$
where $\lambda$ is the penalty parameter, with p different predictors.

@@

To fit the LASSO regression model, we used the `glmnet` package in R.

@@

>>>>

# Load necessary libraries
library(tidyverse) # so many things
library(knitr) # making nice tables
library(caret) # training models & CV
library(leaps) # subset selection
library(corrplot) # visualize correlation
library(rmarkdown) # for output documents
library(rpart) # fitting classification tree
library(rpart.plot) # visualization of classification tree
library(randomForest) # fitting random forest model
library(kernlab) # fitting SVM
library(class) # fitting kNN model
library(glmnet) # fitting LASSO 
library(gbm) # fitting the boosted tree model

@@

  * `caret`: to do the heavy lifting of training and tuning the models
  * `tidyverse`: for all the data reading and wrangling
  * `knitr`: for rmarkdown table outputs
  * `rmarkdown`: for output documents
  * `corrplot`: for correlation structure visualization
  * `leaps`: for subset selection
  * `rpart`: fitting the basic classification tree
  * `rpart.plot`: visualization of the classification tree
  * `randomForest`: fitting random forest models
  * `kernlab`: fitting the support vector machine
  * `class`: fitting the k-nearest neighbor model
  * `glmnet`: fitting the LASSO model
  * `gbm`: fitting the boosted tree model

@@

```{r}
# Create model and get summary output
lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
smre <- lm.sub %>% summary()

@@

# Best subset selection
best.betas <- coef(lm.sub, id = 6) %>% round(2)
```

@@

# Print out regression tree visualization
#tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
#rpart.plot(tree.reg.out, main="Regression Tree")

# Print out classification tree visualization
#tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
#rpart.plot(tree.class.out, main="Classification Tree")

# Print out confusion matrix
#list("RMSE" = tree.reg.rmse,
#     "Class Results" = tree.class.res)
```

@@

# Obtain RMSE and results
rf.reg.pred <- predict(rf.reg, newdata = tst)
rf.reg.rmse <- sqrt(mean((rf.reg.pred - tst$quality)^2))

@@

# Print out results in a list
#list("RMSE" = rf.reg.rmse,
#     "Class Results" = rf.class.res)
```

@@

# Print out results in a list
#list("Results" = svm.class.res,
#     "Confusion Matrix" = svm.class.CM$table)
```

@@

# Obtain RMSE and results
knn.reg.pred <- predict(knn.reg, newdata = tst)
knn.reg.rmse <- sqrt(mean((knn.reg.pred - tst$quality)^2))

@@

# Print out results in a list
#list("RMSE" = knn.reg.rmse,
#     "Class Results" = knn.class.res)
```

@@

# Print out results in a list
#list("Results" = glm.class.res,
#     "Confusion Matrix" = glm.class.CM$table)
```

## LASSO Regression
LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero using a penalty term. This is different from traditional shrinkage methods since the form of the LASSO penalty actually forces some of the coefficients to exactly zero. This helps achieve variable selection and yields sparser models, similar to subset selection. In particular, the LASSO estimate minimizes:
$$
\sum_{i=1}^n\left( y_i - \beta_0 - \sum_{j=1}^p\beta_jx_{ij}\right)^2 + \lambda \sum_{j=1}^p|\beta_j| = RSS + \lambda \sum_{j=1}^p|\beta_j|
$$
where $\lambda$ is the penalty parameter, with p different predictors.

@@

To fit the LASSO regression model, we used the `glmnet` package in R.
```{r}
#### REGRESSION ####
# Fit LASSO to the data
lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))

lasso.reg.pred <- predict(lasso.reg, newdata = tst)
lasso.reg.rmse <- sqrt(mean((lasso.reg.pred - tst$quality)^2))
#lasso.reg.rmse
```

@@

```{r}
#### REGRESSION ####
# Fit the regression boosted tree model 
boost.reg <- train(form.reg, data = trn, method = "gbm",
                trControl = trctrl, verbose = FALSE)

# Obtain RMSE and results
boost.reg.pred <- predict(boost.reg, newdata = tst)
boost.reg.rmse <- sqrt(mean((boost.reg.pred - tst$quality)^2))
#boost.reg.rmse
```
